---
title: "Part 2 Chi squared with Lizard Tails"
author: "Katherine Rosencrance"
date: "11/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(broom)
```

## Read in data
```{r}
lizards <- read_csv(here("data", "lizards.csv"))
```

Get counts of lizards by species
```{r}
lizards %>% 
  count(spp) %>% 
  arrange(-n)
```

## Make a subset 

- UTST lizards
- sites: IBPE (enclosure grassland), RABB (rabbit mesquite site), EAST (east tarbush site)
- tail condition is recorded 

```{r}
utst <- lizards %>% 
  filter(spp == "UTST") %>% 
  filter(site %in% c("IBPE", "RABB", "EAST")) %>% 
  filter(tail %in% c("B", "W"))
```

## find counts and proportions of lizards with broken or whole tails at each site

```{r}
utst %>% 
  count(site, tail)
```

```{r}
tail_counts <- utst %>% 
  janitor::tabyl(site, tail)

tail_proportions <- tail_counts %>% 
  adorn_percentages() %>% 
  janitor::adorn_pct_formatting(digits = 2) %>% 
  janitor::adorn_ns()
  
```

## Convert to contingency table 

use column_to_rowname() function to do that

```{r}
lizard_ct <- tail_counts %>% 
  column_to_rownames(var = "site")
```

## Chi-square

```{r}
lizard_x2 <- chisq.test(lizard_ct)
lizard_x2


lizard_tidy <- tidy(lizard_x2)
```

There is a significant association between site and lizard tail breakage ($\chi$^2^(`r lizard_tidy$parameter`)= `r lizard_tidy$statistic`, p = `r lizard_tidy$p.value`)
